### 2022-01-31 starter_folder-main/outputs/paper/paper.Rmd
Completed visualizations

---
title: "Shooting Occurences in Toronto Have Increased Greatly All Throughout"
subtitle: "My subtitle if needed"
author: 
  - Thomas D'Onofrio
thanks: "Code and data are available at: https://github.com/TDonofrio62/Paper-1."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(knitr)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

(Table \@ref(fig:dataextract))
```{r datextract}

shooting_occurences <- read_csv(here::here("inputs/data/shooting_occurences.csv"))

shooting_occurences %>%
  slice(1,2,18,19,36,37,54,55,72,73,90,91) %>%
  select(occurred_year, geo_division, num_shootings) %>%
  kable(
    caption= "12 Observations from dataset of shooting occurences in Toronto",
    col.names = c("Year", "Geographic Division", "Number of Shootings"),
    booktabs = TRUE,
    linesep = "",
    align = c('c', 'c', 'c')
  )

```

```{r yearlyshootings}
shooting_occurences %>%
  select(occurred_year,num_shootings)%>%
  group_by(occurred_year)%>%
  summarize(number_of_shootings = sum(num_shootings)) %>%
  kable(
    caption= "Number of Shootings in Toronto each year from 2014-2019",
    col.names = c("Year", "Number of Shootings"),
    booktabs = TRUE,
    linesep = "",
    align = c('c', 'c')
  )

```

```{r, fig.cap= "Relationship between number of shootings in each police division and year"}
shooting_occurences %>%
  ggplot(aes(x = occurred_year, y = num_shootings, color = geo_division)) +
  geom_point()+
  theme_bw() +
  labs(x = "Year",
       y = "Number of Shootings",
       color = "Geographic Division"
       )
```

```{r}
shooting_occurences %>%
  ggplot(aes(x = geo_division, y = num_shootings, color = as.character(occurred_year))) +
  geom_point()+
  theme_bw() +
  labs(x = "Geographic Division",
       y = "Number of Shootings",
       color = "Year"
       )+
   scale_color_brewer(palette = "Set1", direction=-1)
```

```{r}
shooting_occurences %>%
  ggplot(aes(x = occurred_year, y = num_shootings, fill = geo_division)) + 
  geom_bar(stat='identity') +
  theme_bw() +
  labs(x = "Year",
       y = "Number of Shootings",
       fill = "Geographic Division"
       )
```

```{r}
shooting_occurences %>%
  ggplot(aes(x = geo_division, y = num_shootings, fill = as.character(occurred_year))) + 
  geom_bar(stat='identity') +
  theme_bw() +
  labs(x = "Geographic Division",
       y = "Number of Shootings",
       fill = "Year"
       )+
  scale_fill_viridis_d()
```

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References